---
title: "Data structures in R"
author: "Andrew Ba Tran"
output:
  html_document:
    toc: yes
    toc_float: yes
description: https://learn.r-journalism.com/en/how_to_use_r/
---

This is from the [first chapter](https://learn.r-journalism.com/en/how_to_use_r/initial_exploration/data-structures/) of [learn.r-journalism.com](https://learn.r-journalism.com/).


## Vectors

A **vector** is a sequence of data elements of the same basic type. The parts that consist of a vector are called **components** or **elements**.

```{r vectors1}
vec1 <- c(1,4,6,8,10)
vec1
```

A vector `vec` is explicitly constructed by the concatenation function `c()`.

```{r vectors2}
vec1[5]
```

Elements in vectors can be addressed by standard `[i]` indexing

```{r vectors3}
vec1[3] <- 12
vec1
```

One of the elements in the array is replaced with a new number. 

```{r vectors4}
vec2 <- seq(from=0, to=1, by=0.25)
vec2
```

This shows another useful way of creating a vector: the `seq()` or sequence function.

```{r vectors5}
sum(vec1)
```

## Data frames

If you're used to working with spreadsheets, then *data frames* will make the most sense to you in R.

This is how to create a data frame from arrays. You don't have to fully understand this at this point-- the data you'll be working with will come pre-structured if you're importing spreadsheets.


```{r df1}
patientID <- c(111, 208, 113, 408)
age <- c(25, 34, 28, 52)
sex <- c(1,2,1,1)
diabetes <- c("Type1", "Type2", "Type1", "Type1")
status <- c(1,2,3,1)

patientdata <- data.frame(patientID, age, sex, diabetes, status)
patientdata
```

But this is what's happening. A set of arrays are being created and a function called `data.frame()` joins them together into a data frame structure.

**How to pull elements from a data frame:**

```{r patientdata}
# a : means "through"
patientdata[1:2]

# So 1:2 means 1 through 2


patientdata[c("diabetes", "status")]

patientdata$age

patientdata[1:2]

patientdata[c(1,3),1:2]

patientdata[2:3, 1:2]
```

```{r df2}
mean(patientdata$age)
```

Instead of using `mean(patientdata[,2])`, you can select the column `age` from the `patientdata` data frame with the `$` sign.


```{r df3}
mean(patientdata[["age"]])
```

Here's an alternative way to refer to the `age` column of the `patientdata` data frame. But you will rarely use this method.

## Lists

Another basic structure in R is a *list*. 

The main advantage of lists is that the "columns" they're not really ordered in columns any more, but are more of a collection of vectors) don't have to be of the same length, unlike matrices and data frames.

Kind of like JSON files are structured.

```{r list1}
g <- "My First List"
h <- c(25, 26, 18, 39)
# The line below is creating a matrix that's 5 rows deep of numbers 1 through(":") 10
j <- matrix(1:10, nrow = 5) 
k <- c("one", "two", "three")
mylist <- list(title = g, ages = h, j, k)   
```

This is how a list would appear in the work space

```{r list2}
names(mylist)
```

How to find out what's in the list

```{r list_extract}
mylist[[2]] 

mylist[["ages"]][[1]]
```

The code above extracts data from the list

```{r list3}
mylist$age + 10
```

How to refer to and use the numbers in the example list

## Functions for working with objects

Let's start with the sample_df dataframe below.

```{r factor1}
# Run the lines of code below
sample_df <- data.frame(id=c(1001,1002,1003,1004), name=c("Steve", "Pam", "Jim", "Dwight"), age=c(26, 65, 15, 7), race=c("White", "Black", "White", "Hispanic"))
sample_df$name <- as.character(sample_df$name)
sample_df
```

`length(x)` - Find out how many things there are in an object or array

```{r length}
length(sample_df$name)
```

`nchar(x)` - If **x** is a string, finds how how many characters there are 

```{r nchar}
sample_df$name[1]
nchar(sample_df$name[1])
```

`dim(x)` - Gives the dimensions of **x**
```{r dim}
dim(sample_df)
```

`ncol(x)` - Counts the number of columns

```{r ncol}
ncol(sample_df)
```

`nrow(x)` - Returns the number of rows of **x**

```{r nrow}
nrow(sample_df)
```

`str(x)` - Returns the structure of **x**

```{r str}
str(sample_df)
```

`summary(x)` - Summarizes the object as understood by R

```{r summary}
summary(sample_df)
```


`View(x)` - A command to open the object to browse in RStudio

```{r view, eval=F}
View(sample_df)
```

`rm(x)` - Removes **x**

```{r rm, error=T}
rm(sample_df)
sample_df
```

## Your turn

Challenge yourself with [these exercises](http://code.r-journalism.com/chapter-1/#section-data-structures-in-r) so you'll retain the knowledge of this section.

Instructions on how to run the exercise app are in the [intro page](http://learn.r-journalism.com/en/how_to_use_r/) to this section.



