---
title: "Chapter 1"
tutorial:
  id: "code.r-journalism/chapter-1"
  version: .85
output:
  learnr::tutorial:
    progressive: true
    theme: lumen
    highlight: espresso
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)

years <- c(301, 978, 101)
burgers <- data.frame(id=c(60006,60007,60008,60009, 60010), name=c("Bob", "Linda", "Louise", "Tina", "Gene"), age=c(45, 44, 12, 13, 11), shirt=c("White", "Red", "Pink", "Blue", "Yellow"))
burgers$shirt<- factor(burgers$shirt)
burgers$id <- factor(burgers$id)
burgers$name <- as.character(burgers$name)

nospace <- function(a) {
  b <- gsub(" ","",a) 
  gsub("\\n", "", b)
}





is_bad_code <- function(user_code, check_code, envir_result) {

  is_null <- is.null(eval(parse(text = user_code)))
  e_rows <- nrow(eval(parse(text = user_code)))
  r_rows <- nrow(eval(parse(text = check_code), envir = envir_result))
  e_cols <- ncol(eval(parse(text = user_code)))
  r_cols <- ncol(eval(parse(text = check_code), envir = envir_result))
  e_type <- typeof(eval(parse(text = user_code)))
  r_type <- typeof(eval(parse(text = check_code), envir=envir_result))
  e_len <- length(eval(parse(text = user_code)))
  r_len <- length(eval(parse(text = check_code)))
  if (is_null!=T){
    if (e_len!=0 & r_len!=0) {
      if (e_type==r_type) {
       if (e_type!="character" & e_type!="double" & e_type!="integer" & e_type!="logical")  {
          if (e_rows==r_rows && e_cols==r_cols) {
            eval(parse(text = user_code)) != eval(parse(text = check_code), envir = envir_result)
          } else {
            TRUE
          }
        } else { eval(parse(text = user_code)) != eval(parse(text = check_code), envir = envir_result) }
      } else {
        TRUE
      }
    } else { TRUE }
  } else { nospace(user_code) != nospace(check_code)}
}


checker <- function(label, user_code, check_code, envir_result, evaluate_result, ...) {
  if (is_bad_code(user_code, check_code, envir_result)) {
    return(list(message = "Code wasn't right!", correct = FALSE, location = "append"))
  } else {
    return(list(message = "Great Job!", correct = TRUE, location = "append"))
  }
}  

tutorial_options(exercise.timelimit = 30, exercise.checker = checker)
#tutorial_options(exercise.checker = checkthat::check_exercise)


```


## Intro to R


### Objects

Assign the number 17 to the object **ub**


```{r filter, exercise=TRUE}
ub

ub
```


```{r filter-check}
ub <- 17

ub
``` 

### Array

Create an array of numbers: 301, 978, and 101.

Assign it to the object "years"

```{r y1, exercise=TRUE}
years #replace this with your code 

years
```


```{r y1-check}
years <- c(301, 978, 101)

years
``` 


### Functions


What's the average of the array of numbers assigned to "years"?

```{r a1, exercise=TRUE}
(years)
```


```{r a1-check}
mean(years)
``` 

### Classes

Take a look at the structure of **burgers**:


```{r st1, exercise=TRUE}

```


```{r st1-check}
str(burgers)
``` 


```{r first_quiz}
quiz(
  question("What kind of class is the variable id?",
    answer("character"),
    answer("number"),
    answer("factor", correct = TRUE),
    answer("date"),
    random_answer_order= TRUE
  ))
```

## Data structures in R


### Pulling a column of data

Consider this data frame **burgers**

```{r burger_show}
burgers
```


How do you refer to the the *shirt* variable/column with **[]**?


```{r v1, exercise=TRUE}
# Add to the line below 

burgers
```



```{r v1-check}
burgers[,4]
``` 


How do you refer to the the *shirt* variable/column with $?

```{r v2, exercise=TRUE}
# Add to the line below 

burgers
```

```{r v2-check}
burgers$shirt
``` 



### Pulling a row of data

Extract entire row for Linda using [].

```{r v4, exercise=TRUE}
# Add to the line below 

burgers
```

```{r v4-check}
burgers[2,]
```   


### Converting data classes

Convert the *id* variable of the **burgers** data frame to numeric.


```{r v3, exercise=TRUE}
# Add to the line below 

burgers$id <- 

burgers$id
class(burgers$id)
```

```{r v3-check}
burgers$id <- as.numeric(as.character(burgers$id))

burgers$id
class(burgers$id)
```   


### Boolean logic

Check if Gene's age is 11.

*Note:* Is the answer the same as above (correct) or is it 1-5 (false)? 


```{r b, exercise=TRUE}
# Modify the line of code below 

age_test <- burgers$age[5]    11

age_test
```

```{r b-check}
age_test <- burgers$age[5] ==   11

age_test
```   
