---
title: "Setting a project folder for class"
description: "Instructions on how to follow along with this class and get the right data"
author: "Andrew Ba Tran"
date: 2018-05-27T21:13:14-05:00
categories: ["R"]
tags: ["R", "projects", "workflow"]
weight: 4
slug: class-project
---

The files for this class will include the raw data, the scripts, and r markdown files.

There are a few options to download them:

1. Download the entire repo off Github
2. Install the package I've created for this class

```{r install, eval=F}
install.packages("devtools")
devtools::install_github("andrewbtran/lrj")
```

3. Follow the directions at the top of each following classes to install

```{r install2, eval=F}
library(lrj)
install_chapter(1)
```

4. Or create the files yourself from scratch the correct folder structure
    * Create a new project in RStudio right now and call it learning-r-journalism
    * Save all your scripts and your data into here for now (ideally with an organized folder structure)

----

Even if the script with the correct exists on this website and in the downloaded files, type as much of the code I describe in the videos and on this site yourself. 

Avoid copying and pasting the code because you’re passively learning.

Do as much manually as you can. This starts the muscle memory and problem solving training.

That being said, if you get stuck, all the scripts and code and proper folder structure is available to download. 
